# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data

```{r}
rawdata<-read.csv("activity.csv") ## orginal data
refineData<-na.omit(rawdata) ## remove NA value data
summary(rawdata)
summary(refineData)
```

## What is mean total number of steps taken per day?

### 1. Make a histogram of the total number of steps taken each day

```{r}
##Total number of steps taken per day
Total.Steps<-tapply(rawdata$steps,rawdata$date,sum)
hist(Total.Steps,10,main="Total number of steps taken per day",xlab="Total Steps")
```

### 2. Calculate and report the mean and median total number of steps taken per day

```{r}
## print mean and median total number of steps taken per day
Total.Steps_mean<-mean(Total.Steps,na.rm=TRUE)
Total.Steps_median<-median(Total.Steps,na.rm=TRUE)
Total.Steps_mean
Total.Steps_median
```

## What is the average daily activity pattern?

### 1. Make a time series plot of the 5-minute interval and the average number of steps taken.
```{r} 
Pattern <- aggregate(refineData$steps, by=list(refineData$interval), FUN=mean)
plot(Pattern[,1],Pattern[,2],type="l",main="Average daily activity pattern",xlab="interval",ylab="average steps")
```

### 2. Which 5-minute interval, on average across all the days in the datase?

```{r} 
Max_interval<-Pattern[which.max(Pattern[,2]),1] ## interval contains the maximum number of steps
Max_interval
```

## Imputing missing values

### 1. Report the total number of missing values in the dataset
```{r}
## the total number of missing values in the dataset 
sum(is.na(rawdata$steps))
```

### 2. Devise a strategy for filling in all of the missing values in the dataset.

- I will use the 5-minute interval data of the average number of steps taken for filling missing values.
- And I will use also KnnImputation method using DmWR package.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
## The method is using the mean as a that 5-minute interval for filling
first_imputing<-rawdata
first_imputing$steps[is.na(first_imputing$steps)]<-Pattern[,2] ## using above pattern data
summary(first_imputing)
##install.packages("DMwR")
library(DMwR)
second_imputing<-knnImputation(rawdata)
summary(second_imputing)
```

### 4. Make a histogram of the total number of steps taken each day and report the mean and median

```{r}
first_imputing.Steps<-tapply(first_imputing$steps,first_imputing$date,sum)
first_imputing.Steps_mean<-mean(first_imputing.Steps,na.rm=TRUE)
first_imputing.Steps_median<-median(first_imputing.Steps,na.rm=TRUE)
second_imputing.Steps<-tapply(second_imputing$steps,second_imputing$date,sum)
second_imputing.Steps_mean<-mean(second_imputing.Steps,na.rm=TRUE)
second_imputing.Steps_median<-median(second_imputing.Steps,na.rm=TRUE)
par(mfrow=c(1,2))
hist(first_imputing.Steps,10,main="Data by imputing the mean",xlab="Total Steps")
hist(second_imputing.Steps,10,main="Data by kNN imputation ",xlab="Total Steps")
```

========================================================================================================
```{r}
paste("The means of original data and impuing data are",Total.Steps_mean,",",first_imputing.Steps_mean,"and",second_imputing.Steps_mean)
paste("The medians of original data and impuing data are",Total.Steps_median,",",first_imputing.Steps_median,"and",second_imputing.Steps_median)
```

## Are there differences in activity patterns between weekdays and weekends?


### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend"

```{r}
## Create a new factor variable in the dataset with two levels 
## ¡°weekday¡± and ¡°weekend¡± 
Sys.setlocale("LC_TIME", "English")
week<-ifelse(weekdays(as.Date(second_imputing$date))=="Sunday"|weekdays(as.Date(second_imputing$date))=="Saturday","weekend","weekday")
week<-as.factor(week)
newdata<-cbind(second_imputing,week)
summary(newdata)
```

### 2. Make a panel plot containing a time series plot.

```{r}
newPattern <- aggregate(newdata$steps, by=list(newdata$interval,newdata$week), FUN=mean)
colnames(newPattern)<-c("interval","week_type","activity")
weekend_data<-subset(newPattern,newPattern$week_type=="weekend")
weekday_data<-subset(newPattern,newPattern$week_type=="weekday")
par(mfrow=c(2,1),mar=c(4,4,2,1))
plot(weekend_data$interval,weekend_data$activity,type="l", xlab="", ylab="no of steps", main="weekend",col="blue")
title(main="#steps, split on weekend/weekdays", outer=TRUE)    
plot(weekday_data$interval,weekday_data$activity,type="l", xlab="5 min interval", ylab="no of steps", main="weekday",col="blue")
```

